Hierarchical Time Series Forecasting in Emergency Medical Services

### 2022-11-18 manuscript/03-experiment.qmd
sec 5 and 6

**Poisson Regression using tscount:** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents. Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive. This model captures the short range serial dependence by including three autoregressive terms, in addition to the same covariates that were used in the GLM model. To distinguish this from the previous GLM model, we will refer to this model as "tscount".

**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and tscount models to form a mixture distribution [@combinations].

>>>>

**Poisson Regression using tscount (TSGLM):** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents. Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive. This model captures the short range serial dependence by including three autoregressive terms, in addition to the same covariates that were used in the GLM model. To distinguish this from the previous GLM model, we will refer to this model as "TSGLM".

**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models to form a mixture distribution [@combinations].

### 2022-11-18 manuscript/05-results.qmd
sec 5 and 6

# Results and discussion {#sec-result}

## Point forecast

## Probabilistic forecast

```{r}
#| label: fig-accuracy
#| fig-width: 8
#| fig-height: 5
#| out.width: "100%"
#| fig-cap: "Average accuracy by week for 12 weeks. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
mase <- readr::read_rds(here::here("results/mase.rds"))
crps <- readr::read_rds(here::here("results/crps.rds")) |>
  pivot_wider(names_from = model, values_from = crps) |>
  group_by(method, h, series) |>
  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
  ungroup() |>
  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")

@@

# Plot of average accuracy vs week for each method for Total
acc_summary <- accuracy |>
  filter(
    series %in% c("Total", "Control areas", "Health boards"),
    method == "mint", model != "qcomb"
  ) |>
  mutate(
    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
    week = factor(trunc((h - 1) / 7) + 1)
  ) |>
  group_by(week, model, measure, series) |>
  summarise(accuracy = mean(accuracy), .groups = "drop")

acc_summary |>
  ggplot(aes(x = week, y = accuracy, group = model)) +
  geom_line(aes(col = model)) +
  facet_grid(measure ~ series, scales = "free_y") +
  labs(y = "Average accuracy", x = "Week ahead")
```

>>>>

# Results and discussion {#sec-results}

```{r}
#| include: false
rmsse <- readr::read_rds(here::here("results/rmsse.rds")) |> 
  mutate(msse = rmsse^2)
mase <- readr::read_rds(here::here("results/mase.rds"))
crps <- readr::read_rds(here::here("results/crps.rds")) |>
  pivot_wider(names_from = model, values_from = crps) |>
  group_by(method, h, series) |>
  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
  ungroup()
```

In this section, we compare the forecasting performance of the Naive, ETS, GLM and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods. We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches, along with bottom up forecasting. However, they are not reported here because their accuracy is outperformed by MinT. We should also note that forecasts, and consequently their corresponding errors, are generated for the entire hierarchy and they could be reported at any level, if required. But to save space, we have reported only the top level (Total), the bottom level, and the levels corresponding to Control areas and Health boards. The latter are chosen because this is where decision-making takes place, so these forecasts are the most important.

The overall forecasting performance is reported in @tbl-result, in which the average forecast accuracy over horizons 43--84 days (corresponding to the planning horizon) is presented per model, method and the hierarchical level. Reported forecast accuracy are averaged across all forecast horizons, rolling origins and series at each level. @tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series. Point forecast performance are reported using MASE and MSSE in @tbl-result-1  and @tbl-result-2, respectively. Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3. The bold entries in each table identify a combination of method and model that performs best for the corresponding level (i.e. each column), based on the smallest values of accuracy measures.

@tbl-result-1 and 2b show that forecast reconciliation (i.e. MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board. However, it is does not result in accuracy improvement at the bottom level series, for which base forecasts are more accurate. This might be due to the noisy structure of time series at the bottom level, and possibly due to very different patterns in the aggregated series. It is also clear from @tbl-result-1 that the ensemble method improves forecast accuracy at total, control area and health board. However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.

```{r}
#| label: tbl-result
#| tbl-cap: "Average forecast perfomance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data. The best approach is highlighted in bold."
#| tbl-subcap:
#|   - "Point forecast accuracy using MASE"
#|   - "Point forecast accuracy using MSSE"
#|   - "Probabilistic forecast accuracy using CRPS"
#| layout-nrow: 3
#Set minimum in column to bold
set_to_bold <- function(table) {
  for(i in 3:6) {
    best <- which.min(table[[i]])
    table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
    table[[i]][best] <- cell_spec(table[[i]][best], "latex", bold=TRUE)
  }
  return(table)
}

msse1 <- rmsse |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42
  ) |>
  mutate(model = factor(model,
    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
  ) |>
  group_by(method, model, series) |>
  summarise(msse = mean(msse), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = msse) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
mase1 <- mase |>
  filter(
    method %in% c("mint", "base"),
    model != "qcomb",
    h > 42
  ) |>
  mutate(
    model = factor(model,
      levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
      labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(mase = mean(mase), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = mase) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()
crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
  filter(method %in% c("mint", "base"),
         model != "qcomb",
         h > 42
  ) |>
  mutate(model = factor(model,
    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
  ) |>
  group_by(method, model, series) |>
  summarise(crps = mean(crps), .groups = "drop") |>
  pivot_wider(names_from = series, values_from = crps) |>
  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
  set_to_bold()

kbl(mase1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
  add_header_above(c(" " = 2, "MASE" = 4)) |>
  kable_styling(latex_options = c("hold_position"))

kbl(msse1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
  add_header_above(c(" " = 2, "MSSE" = 4)) |>
  kable_styling(latex_options = c("hold_position"))

kbl(crps1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
  add_header_above(c(" " = 2, "CRPS" = 4)) |>
  kable_styling(latex_options = c("hold_position"))
```

@tbl-result-3 presents the accuracy of the forecast distributions measures by CRPS, which considers both forecasting reliability and interval sharpness. The smaller the value of CRPS, the better the comprehensive performance. We observe that forecast reconciliation results in forecast improvement, regardless of the hierarchical level. The ensemble method is also more accurate for higher levels, but ETS performs slightly better at the bottom level. While these results show that forecast reconciliation does not improve *point* forecasts at the bottom level, @tbl-result-3 indicates that it generates more accurate *distributional* forecasts than the base method. This is probably due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.

Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts, improves upon the base (unreconciled) forecasts all levels except the bottom level series. But even there, forecast reconciliation using MinT improves accuracy in the tails of the distribution.

In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy. The figure focuses on the hierarchical levels important for decision-making including total, control area and health board; however the accuracy could be calculated for any level. We only illustrate the results of the MinT method, given its strong performance described in @tbl-result. For illustration purposes, we report the average weekly forecast accuracy instead of the daily forecast horizon, as this reduces the visual noise in the figure. Thus the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$). The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision making. For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation. It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a benchmark for both point and probabilistic forecasts.

```{r}
#| label: fig-accuracy
#| fig-width: 8
#| fig-height: 5
#| out.width: "100%"
#| fig-cap: "Average accuracy by week for 12 weeks using MinT reconciliation. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
#| fig-pos: "H"
crps <- crps |>
  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")

@@

# Plot of average accuracy vs week for each method for Total
acc_summary <- accuracy |>
  filter(
    series %in% c("Total", "Control areas", "Health boards"),
    method == "mint", model != "qcomb"
  ) |>
  mutate(
    measure = factor(measure, levels = c("mase","msse","crps"), labels = c("MASE","MSSE","CRPS")),
    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
    week = factor(trunc((h - 1) / 7) + 1)
  ) |>
  group_by(week, model, measure, series) |>
  summarise(accuracy = mean(accuracy), .groups = "drop")

acc_summary |>
  ggplot(aes(x = week, y = accuracy, group = model, col = model)) +
  geom_line() +
  geom_point(size = .5) +
  facet_grid(measure ~ series, scales = "free_y") +
  labs(y = "Average accuracy", x = "Week ahead") +
  ggthemes::scale_color_colorblind() +
  ggthemes::theme_few()
```

### 2022-11-18 manuscript/06-conclusion.qmd
sec 5 and 6

Forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures. For example, looking at time series of arrival calls in a clinical desk service, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest. Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Alternative structure may arise when attributes of interest are crossed rather than nested. For example, the total demand could be disaggrgated by priority (e.g Red, Amber, Green) or nature of incidents. It is also natural to have a mixed structures, for example the total demand could be disaggregated by priority and also by health board.

Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecast for each time series independently. This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes. Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks. In this paper, we investigate the application of hierarchical forecasting methods for producing a probabilistic forecast of daily incident attended up to 84 days ahead, using different forecasting methods.

Our results indicates that forecast reconciliation can not only contribute to a more coordinated efforts through producing coherent forecasts when it comes to using forecasting to inform decisions in EMS, but also it can result in forecast accuracy improvement. Our proposed forecasting models combined with reconciliation approaches outperform the empirical distribution benchmark. We show that a substantial forecast improvement can be achieved at higher levels of hierarchical and grouped structure time series by applying forecast reconciliation methods. When a point forecast is of interest at the bottom level of series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns. However, forecast reconciliation may result in more forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e. point forecast). Producing consistent forecast are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons. Therefore, we recommend that forecast reconciliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted. Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individualy, works remarkably well for our mixed hierarchical & grouped structure.

>>>>

Forecasting problems at Emergency Medical Services often have inherent hierarchical and grouped structures. For example, looking at time series of arrival calls in a clinical desk service, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest. Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Alternative structures may arise when attributes of interest are crossed rather than nested. For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by nature of incidents. It is also natural to have a mixed structures, for example the total demand could be disaggregated by priority and also by health board.

Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecasts for each time series independently. This practice may lead to a lack of coordination and possibly undesirable and conflicting outcomes. Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks. In this paper, we investigate the application of hierarchical forecasting methods for producing probabilistic forecasts of daily incidents attended up to 84 days ahead, using different forecasting methods.

Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to decision making through producing coherent forecasts, but also it can result in forecast accuracy improvement. Our proposed forecasting models, combined with reconciliation approaches, outperform the empirical distribution benchmark. We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods. When a point forecast is of interest at the bottom level of series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns. However, forecast reconciliation may result in more forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e. point forecast). Producing consistent forecast are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons. Therefore, we recommend that forecast reconciliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted. Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individually, works remarkably well for our mixed hierarchical & grouped structure.
